Eagle: multi-locus association mapping on a genome-wide scale made routine

### 2019-07-24 BioinformaticsRevison/bioinformaticspaper.tex
Revision working through Review 1 changes

What has not changed is that it remains common practice to analyse genome-wide association study (GWAS) data on a locus-by-locus basis. This is despite there being several significant problems with analysing data in this way. 
First, \textcolor{red}{ for each SNP, a hypothesis test is performed. The null hypothesis is that  there is no association between the SNP and trait. 
The alternative is that the SNP is in association with the trait. It is straight forward to guard against wrongly rejecting the null 
hypothesis (or making a type 1 error) if only a single hypothesis test is being performed. However, the analysis of GWAS data 
with locus-by-locus methods necessitates  conducting a large number of correlated hypothesis tests, simultaneously. 
 This leads to an increased risk of type 1 errors.  To deal with this challenge, many different solutions have been offered 
 \citep{storey2003statistical, li2005adjusting, de2005efficiency}}.  Second, 
the aim of association mapping is to identify regions of the genome that house genes that are influencing a trait. 
The identification of these regions from these analyses is not always straightforward. GWAS results are reported, typically, via Manhattan plots 
that plot the -log$_{10}$ of the $p$ value for each locus against the map position of the locus. The $p$ value is obtained from the hypothesis test. 
The location of peaks in this plot identify genomic 
regions of interest. Inferring the exact number of regions though can be difficult if the peaks are not well separated. 
 Third, many of the traits whose genetic secrets we are trying to discover are complex. There will be multiple SNPs in linkage disequilibrium with genes that are influencing the trait. Yet, a locus-by-locus mapping approach only assesses the evidence for association between a single marker locus and trait.

@@

Eagle is a method for multi-locus association mapping on a genome-wide scale. It is based on linear mixed models. It differs from most other single- and multi-locus association mapping methods in that Eagle treats association mapping as a model selection problem (ball2001bayesian,broman2002model,yi2005bayesian). 
The "best" model is found via forward selection. It makes use of a modified form of the Bayesian information criterion, BIC, for model selection.

Eagle does not suffer from multiple testing issues. A "best" model is built iteratively. At each iteration, a hypothesis test is performed. 
Only a small number of iterations are needed in building the "best" model.  In contrast,  for single-locus methods, multiple testing is an 
issue because each SNP is assessed separately, culminating in a large number of hypothesis tests being performed. 
Eagle also 
reports as its findings only those SNPs that are in strongest linkage disequilibrium with  the genes influencing a trait. 
The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
quantitative trait loci in experimental crosses \citep{verbyla2007analysis}.

@@

\section{Discussion/Conclusion}
Eagle is a new linear mixed model based method (and R package) for multi-locus association mapping. It advances the state of association mapping in several ways. 
First, its computational footprint is much smaller than other multi-locus implementations. Eagle makes multi-locus analysis 
practical, even when the datasets are large. Second, the results from
 Eagle are immediately interpretable. They are the set of SNPs in strongest association with the trait where 
each SNP identifies a separate genomic region of interest. Third, it treats association mapping as a model selection problem, avoiding 
 the need for significance thresholds. 
 As we saw in the simulation study, Eagle has considerably higher power than single-locus methods but is comparable in run time.
Also, when analysing the mouse data, Eagle found more than double the SNP-trait associations than 
with single-locus association mapping, the method of choice. Furthermore, these extra findings were all true.

>>>>

What has not changed is that it remains common practice to analyse genome-wide association study (GWAS) data on a locus-by-locus basis. This is despite there being several significant problems with analysing data in this way. 
First, \textcolor{red}{ for each SNP, a hypothesis test is performed. The null hypothesis is that  there is no association between the SNP and trait. 
The alternative is that the SNP is in association with the trait. It is straight forward to guard against wrongly rejecting the null 
hypothesis (or making a type 1 error) if only a single hypothesis test is being performed. However, the analysis of GWAS data 
with locus-by-locus methods necessitates  conducting a large number of correlated hypothesis tests, simultaneously. 
 This leads to an increased risk of type 1 errors.  To deal with this challenge, many different solutions have been offered 
 \citep{storey2003statistical, li2005adjusting, de2005efficiency}.  Second, 
the aim of association mapping is to identify regions of the genome that house genes that are influencing a trait. 
The identification of these regions from these analyses is not always straightforward. GWAS results are reported, typically, via Manhattan plots 
that plot the -log$_{10}$ of the $p$ value for each locus against the map position of the locus. The $p$ value is obtained from the hypothesis test. 
The location of peaks in this plot identify genomic 
regions of interest. Inferring the exact number of regions though can be difficult if the peaks are not well separated. }
 Third, many of the traits whose genetic secrets we are trying to discover are complex. There will be multiple SNPs in linkage disequilibrium with genes that are influencing the trait. Yet, a locus-by-locus mapping approach only assesses the evidence for association between a single marker locus and trait.

@@

\textcolor{red}{
Eagle is a method for multi-locus association mapping on a genome-wide scale. It is based on linear mixed models. It differs from most other single- and multi-locus association mapping methods in that Eagle treats association mapping as a model selection problem \citep{ball2001bayesian,broman2002model,yi2005bayesian}. 
The "best" model is found via forward selection. It makes use of a modified form of the Bayesian information criterion, BIC, for model selection.
A "best" model is built iteratively. At each iteration, a hypothesis test is performed. 
Only a small number of iterations are needed in building the "best" model.  Consequently, 
Eagle does not suffer from multiple testing issues.
In contrast,  for single-locus methods, multiple testing is an 
issue because each SNP is assessed separately, culminating in the need for a large number of hypothesis tests to be  performed. }
Eagle 
reports as its findings only those SNPs that are in strongest linkage disequilibrium with  the genes influencing a trait. 
The methodological foundation for Eagle comes from a whole-genome linkage analysis method that was developed for mapping 
quantitative trait loci in experimental crosses \citep{verbyla2007analysis}.

@@

\section{Discussion/Conclusion}
Eagle is a new linear mixed model based method (and R package) for multi-locus association mapping. It advances the state of association mapping in several ways. 
First, its computational footprint is much smaller than other multi-locus implementations. Eagle makes multi-locus analysis 
practical, even when the datasets are large. Second, the results from
 Eagle are immediately interpretable. They are the set of SNPs in strongest association with the trait where 
each SNP identifies a separate genomic region of interest. Third, it treats association mapping as a model selection problem, avoiding 
 multiple testing issues. 
 As we saw in the simulation study, Eagle has considerably higher power than single-locus methods but is comparable in run time.
Also, when analysing the mouse data, Eagle found more than double the SNP-trait associations than 
with single-locus association mapping, the method of choice. Furthermore, these extra findings were all true.
